Computational design of RNA-based oscillatory circuits

### 2015-09-15 MScWriteup.tex
finalising the report now

@@

\subsection{The sRNA regulatory system}
\label{The sRNA regulatory system}
In bacteria, one mechanism by which gene expression is naturally regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA (Fig. \ref{RBS}). This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA. 
\setcounter{figure}{1}
\begin{figure*}[t!]
\centering
\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS. In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$. These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E. coli} discussed. These chemical repressors are themselves disabled by two chemicals, aTC and IPTG. In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production. We see a `double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA). Thus the presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.  Image reproduced from \cite{Rodrigo2012}.}
\label{ANDGate}
\end{figure*}
\setcounter{figure}{0}

@@

In this system, transcription of the designed sRNA and mRNA are placed under the control of promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}. These are in turn controlled by two transcriptional repressors, TetR and LacI respectively, which are naturally present in the strain of \textit{E. coli} used. These repressors disable the promoter regions, and so by default transcription of both RNA's is turned off, and no protein is produced. 
These repressors can themselves be disabled by the presence of two chemicals, aTc and IPTG, which can be introduced externally into the cell (Fig. \ref{ANDGate}).  So transcription of the two RNAs is indirectly controlled by the presence of two chemicals - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced. If only one is present, the AND gate remains off, either because there is no mRNA to be translated into protein, or because the mRNA is self repressed. But when both are present, the conformational change discussed above occurs, and protein is produced.

@@

\begin{figure*}
    \begin{subfigure}[c]{0.49\textwidth}
    \centering
    \includegraphics[scale = 0.24,clip = true, trim = 115 0 60 0]{13_9_bestPlot}
        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
                \label{InitialResults_13_9_best}
    \end{subfigure}
    \begin{subfigure}[c]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24,clip = true, trim = 100 0 70 0]{14_7_bestPlot}
        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
        \label{InitialResults_14_7_best}
    \end{subfigure}
        \begin{subfigure}[h]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24, clip = true, trim = 140 0 125 0]{13_9_hist}
        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
        \label{InitialResults_13_9_hist}
    \end{subfigure}
    \begin{subfigure}[c]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24,clip = true, trim = 140 0 125 0]{14_7_hist}
        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
        \label{InitialResults_14_7_hist}
    \end{subfigure}
    \caption{Parameter estimates and model predictions for the \textit{13\_9} and \textit{14\_7} datasets. Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response. The forcing curve's height is schematic - aTc concentration is switched between off, and a level which saturates the cell's response.}
\label{InitialResults}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[scale = 0.23,clip = true, trim = 90 0 60 0]{Sensitivity_scaled_simplified}
\caption{Using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, normalised sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.}
\label{Sensitivity_scaled_simplified}
\end{figure}

@@

 \begin{figure}[t!]	
	\begin{subfigure}[h]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24 ,clip = true, trim = 55 0 0 0]{Likelihood_profile_rough}
        \caption{ A basin of low error values in $\mu - F$ space. Note the colour scaling maps all values above $20$ to the maximum colour.}
        \label{Likelihood_profile_rough} 
    \end{subfigure}
    	\begin{subfigure}[h]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24 ,clip = true, trim = 55 0 0 0]{Likelihood_profile_fine}
        \caption{Zoomed in on the bottom of the basin in Fig.\ref{Likelihood_profile_rough}. Note the colour scaling maps all values above $5.7$ to the maximum colour.}
        \label{Likelihood_profile_fine} 
    \end{subfigure}
    \caption{Error function plotted as a function of $\mu$ and $F$, $\Theta = 5400$. Note the many disconnected local minima in the bottom of a basin of low error function values.} 
\label{SimplifiedModelLikelihood}   
\end{figure}

>>>>

\usepackage{letltxmacro}
\LetLtxMacro{\originaleqref}{\eqref}
\renewcommand{\eqref}{eq.~\originaleqref}

@@

\subsection{The sRNA regulatory system}
\label{The sRNA regulatory system}
In bacteria, one mechanism by which gene expression is naturally regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA (Fig. \ref{RBS}). This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA. 
\setcounter{figure}{1}
\begin{figure*}[t!]
\centering
\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS. In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$. These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E. coli} discussed. These chemical repressors are themselves disabled by two chemicals, aTC and IPTG. In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production. We see a `double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA). Thus the presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.  Image reproduced from \cite{Rodrigo2012}.}
\label{ANDGate}
\end{figure*}
\setcounter{figure}{0}

@@

In this system, transcription of the designed sRNA and mRNA are placed under the control of promoters, $P_{\mathrm{LtetO-1}}$ and $P_\mathrm{LlacO-1}$ \cite{Lutz1997}. These are in turn controlled by two transcriptional repressors, TetR and LacI respectively, which are naturally present in the strain of \textit{E. coli} used. These repressors disable the promoters, and so by default transcription of both RNA's is turned off, and no protein is produced. 
These repressors can themselves be disabled by the presence of two chemicals, aTc and IPTG, which can be introduced externally into the cell (Fig. \ref{ANDGate}).  So transcription of the two RNAs is indirectly controlled by the presence of two chemicals - if neither is present, sRNA and mRNA transcription is repressed, and no protein is produced. If only one is present, the AND gate remains off, either because there is no mRNA to be translated into protein, or because the mRNA is self repressed. But when both are present, the conformational change discussed above occurs, and protein is produced.

@@

\begin{figure*}
    \begin{subfigure}[c]{0.49\textwidth}
    \centering
    \includegraphics[scale = 0.24,clip = true, trim = 115 0 60 0]{13_9_bestPlot}
        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{13\_9} dataset.}
                \label{InitialResults_13_9_best}
    \end{subfigure}
    \begin{subfigure}[c]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24,clip = true, trim = 100 0 70 0]{14_7_bestPlot}
        \caption{Model prediction, using the parameter set with the smallest error value of the initial 200 found, for the \textit{14\_7} dataset.}
        \label{InitialResults_14_7_best}
    \end{subfigure}
        \begin{subfigure}[h]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24, clip = true, trim = 140 0 125 0]{13_9_hist}
        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{13\_9} dataset.}
        \label{InitialResults_13_9_hist}
    \end{subfigure}
    \begin{subfigure}[c]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24,clip = true, trim = 140 0 125 0]{14_7_hist}
        \caption{Histogram of estimated parameter values, found from 200 runs of the CMA-ES algorithm. Fitted to the \textit{14\_7} dataset. }
        \label{InitialResults_14_7_hist}
    \end{subfigure}
    \caption{Parameter estimates and model predictions for the \textit{13\_9} and \textit{14\_7} datasets. Note that in the model predictions, aTc forcing is shown - IPTG concentration is constant at a level which saturates the cell's response. The forcing curve's height is schematic - aTc concentration is switched between off and on (where by `on' we mean a level which saturates the cell's response).}
\label{InitialResults}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[scale = 0.23,clip = true, trim = 80 0 60 0]{Sensitivity_scaled_simplified}
\caption{Using the simplified model \eqref{eq:p_simplified} - \eqref{eq:z_simplified}, normalised sensitivity coefficients $S_{ij}$ evaluated about a set of estimated parameters from the \textit{13\_9} dataset, for a single oscillation.}
\label{Sensitivity_scaled_simplified}
\end{figure}

@@

 \begin{figure}[t!]	
	\begin{subfigure}[h]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24 ,clip = true, trim = 55 0 0 0]{Likelihood_profile_rough}
        \caption{ A basin of low error values in $\mu - F$ space. Note the colour scaling maps all values above $20$ to the maximum colour.}
        \label{Likelihood_profile_rough} 
    \end{subfigure}
    	\begin{subfigure}[h]{0.49\textwidth}
    \centering
        \includegraphics[scale = 0.24 ,clip = true, trim = 55 0 0 0]{Likelihood_profile_fine}
        \caption{Zoomed in on the bottom of the basin in Fig.\ref{Likelihood_profile_rough}. Note the colour scaling maps all values above $5.7$ to the maximum colour.}
        \label{Likelihood_profile_fine} 
    \end{subfigure}
    \caption{Error function plotted as a function of $\mu$ and $F$, $\Theta = 5400$ (the modal value of $\Theta$ in Fig. \ref{13_9_hist_simplified}). Note the many disconnected local minima in the bottom of a basin of low error function values.} 
\label{SimplifiedModelLikelihood}   
\end{figure}
